---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo="TRUE"}
library(dplyr)

activity <- read.csv("activity.csv")
```
## Total Number of Steps Taken Per Day

```{r, echo="TRUE"}
total_daily_steps <- aggregate(data =activity, steps ~ date, FUN = sum, na.rm=TRUE)

hist(total_daily_steps$steps, main = "Steps per Day", xlab = "Steps", col = "orange")
```


## What is mean total number of steps taken per day?
```{r, echo="TRUE"}
print(paste("Mean Steps Overall: ",mean(total_daily_steps$steps, na.rm = TRUE)))
print(paste("Median Steps Overall: ", median(total_daily_steps$steps, na.rm = TRUE)))
mean_steps <- aggregate(data =activity, steps ~ date, FUN = mean, na.rm=TRUE)
hist(mean_steps$steps, main = "Average Steps per Day", col = "blue", xlab = "Number of Steps")
median_steps <-aggregate(data =activity, steps ~ date, FUN = median, na.rm=TRUE)

```


###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max_steps <- max(activity$steps, na.rm=TRUE)
activity[which(activity$steps == max_steps),]

```

## Inputing missing values

### Number of missing values

```{r}
sum(is.na(activity$steps))
```

```{r}
full_activity <- activity[1:3]
full_activity[] <- lapply(full_activity, function(x) ifelse(is.na(x), mean(x, na.rm = TRUE), x))
```

### Total steps each day

```{r}
total_full_steps <- aggregate(data =full_activity, steps ~ date, FUN = sum, na.rm=TRUE)
hist(total_full_steps$steps, xlab = "Steps Each Day", main="Total Steps Each Day", col = "green")
```
### Mean steps in revised data set

```{r}
aggregate(data =full_activity, steps ~ date, FUN = mean, na.rm=TRUE)
```

### Median steps in revised data set

```{r}
aggregate(data =full_activity, steps ~ date, FUN = median, na.rm=TRUE)
```
###What is the impact of imputing missing data on the estimates of the total daily number of steps?
The mean and median values both increase.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lattice)

mean_steps$weekend <- ifelse(weekdays(as.Date(mean_steps$date)) %in% c("Saturday","Sunday"), "weekend", "weekday")

panel.smoother <- function(x, y) {
  panel.xyplot(x, y) # show points 
  panel.loess(x, y)  # show smoothed line 
}

with(mean_steps, {
  xyplot(steps ~as.Date(date, "%Y-%m-%d") |factor(weekend), 
         type = "l",
         xlab = "Date", ylab = "Number of Steps", 
         main = "Average Steps per Day on Weekdays/Weekends")    
})
```